---
title: "HW3"
author: "Craig Alder"
date: "February 15, 2017"
output: html_document
---



You are now the data slave to the principal investigator Dr. Vinca Monster. Dr. M is in the Grape Program at State U, and you are just a poor graduate student trying to get your degree. Dr. M is interested in wine preferences and the influences of physico-chemical properties on preferences. Her laboratory has gathered an extensive dataset on Portugese white varietals.


You will find the white_wines.csv file and its description in my github repo (https://github.com/vhertzb/Regression-1)[https://github.com/vhertzb/Regression-1].

Please use the techniques you have learned in the last two classes, specifically exploratory data analysis and linear regression, to determine association of the wine properties on preference. 

Prepare a report for presentation at the next Monster lab meeting about this dataset.

Rubric:

Exploration (summary statistics (the m's), univariate graphs, multivariate graphs)
Regression (Models explored, diagnostics completed, final model choice, justification)

Please include a concluding paragraph (or two) about the implications of your findings.

```{r}
#load up necessary packages
library(HistData)
library(car)
library(stargazer)
```

```{r}
#read in dataset
library(readr)
White_wines <- read_csv("~/Documents/Big Data Class/N741 Data Wrangling/InClass2.1/Regression-1/White_wines.csv")

#Why did I have to do below?
names(White_wines) <- make.names(names(White_wines))
```

```{r}
# see what is in the dataset
summary(White_wines)
```
It doesn't seem as if there is any unaccounted missing data unaccounted.

```{r}
scatterplotMatrix(~quality+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines)

scatterplotMatrix(~quality+pH+density+free.sulfur.dioxide+chlorides,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines)

scatterplotMatrix(~quality+citric.acid+fixed.acidity+alcohol,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines)
```

Free sulfur dioxide, fixed acidity, and density seem to have abnormal residual distributions, so I logged them below.

```{r}
White_wines$free.sulf.diox.log <- with(White_wines, 
  log2(free.sulfur.dioxide))
```

```{r}
White_wines$fixed.acidity.log <- with(White_wines, log2(fixed.acidity))
```

```{r}
White_wines$density.log <- with(White_wines, log2(density))
```

```{r}
scatterplotMatrix(~quality+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines)

scatterplotMatrix(~quality+pH+density.log+free.sulf.diox.log+chlorides,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines)

scatterplotMatrix(~quality+citric.acid+fixed.acidity.log+alcohol,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines)
```

Those transformations don't seem to help much visually, but I will try my model with and without them logged to check.

```{r}
cor(White_wines[,c("quality","alcohol","chlorides","citric.acid","density.log","fixed.acidity.log","free.sulf.diox.log","pH","residual.sugar","sulphates","total.sulfur.dioxide","volatile.acidity")], use="complete")
```
The variable most correlated with quality is alcohol, so I will use that as my primary independent variable. Density is highly correlated with alcohol (r=-.78) and residual sugar (r=.84),and free sulfur dioxide less so with total sulfur dioxide (r=.59), so I will consider whether these variables all need to be in the model.


Below is my first regression, with all potential independent variables included.
```{r}
RegModel.1 <- 
  lm(quality~alcohol+chlorides+citric.acid+density.log+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines)
summary(RegModel.1)
```
**Interpretation:**
For each 1-unit increase in alcohol (I am guessing 1 percent alcohol content), the rating of quality increases by 0.198 on a 7 point scale, holding all other variables (different qualities of the wine) constant. This is significant at p<.001.

This interpretation could be extended to any of the other independent variables. For example, a 1 unit increase of chlorides is associated with a .404 decrease in rating of wine quality, all other independent variables held constant; however, the p-value, .65, is not significant.

```{r}
RegModel.1.2 <- 
  lm(quality~alcohol+chlorides+citric.acid+density+fixed.acidity+free.sulfur.dioxide+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines)
summary(RegModel.1.2)
```
This is the same model without density, fixed acidity, and free sulfur dioxide logged. The R-squared is noticeably reduced by about .02, so I'll keep those variables logged.

```{r}
RegModel.2 <- 
  lm(quality~alcohol+density.log+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines)
summary(RegModel.2)
```

Below, I try model 1 without density, recalling that density is highly correlated with two other variables and has a high standard error in the previous models.
```{r}
RegModel.3 <- 
  lm(quality~alcohol+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines)
summary(RegModel.3)
```
**Interpretation:**
With density dropped, there are no anomolous standard errors. R-squared decreases slightly, but not enough to be practically significant.

```{r results="asis"}
# compare the results of the two regression models
stargazer(RegModel.1,RegModel.2,RegModel.3,title="Comparison of 3 Regression outputs",type="html",align=TRUE)
```

```{r}
# diagnostics using residual plots
residualPlots(RegModel.1)
residualPlots(RegModel.2)
residualPlots(RegModel.3)
```

```{r}
White_wines$tot.sulf.diox.log <- with(White_wines, 
  log2(total.sulfur.dioxide))
```

```{r}

#added variable plots
avPlots(RegModel.3, id.n=3, id.cex=0.7)
#id.n - identify n most influential observations
#id.cex - controls the size of the dot
```

```{r}

# run the qq-plot
qqPlot(RegModel.3, id.n=3)
# here, id.n identifies the n observations with the largest residuals in absolute value
```

Are there any outliers?

```{r}
 
#run Bonferroni test for outliers
outlierTest(RegModel.3)
```

Are there any points that are of high influence?

```{r}

#identify highly influential points
influenceIndexPlot(RegModel.3, id.n=3)
```

NB. If there are points that are a) outliers AND b) highly influential, these have potential to change the inference. You should consider removing them.

How do we make heads or tails out of the plots above? One way is with an influence plot.
```{r}
#make influence plot
influencePlot(RegModel.3, id.n=3)
```

Another diagnostic is to test for heteroskedasticity (i.e., the variance of the error term is not constant).

```{r}
#test for heteroskedasticity
ncvTest(RegModel.3)
```

We also want to look for multicollinearity, that is are some of our independent variables highly correlated. We do this by looking at the Variance Inflation Factor (VIF). A GVIF > 4 suggests collinearity.

```{r}
vif(RegModel.3)
```