---
title: "HW3"
author: "Craig Alder"
date: "February 15, 2017"
output: html_document
---



You are now the data slave to the principal investigator Dr. Vinca Monster. Dr. M is in the Grape Program at State U, and you are just a poor graduate student trying to get your degree. Dr. M is interested in wine preferences and the influences of physico-chemical properties on preferences. Her laboratory has gathered an extensive dataset on Portugese white varietals.


You will find the white_wines.csv file and its description in my github repo (https://github.com/vhertzb/Regression-1)[https://github.com/vhertzb/Regression-1].

Please use the techniques you have learned in the last two classes, specifically exploratory data analysis and linear regression, to determine association of the wine properties on preference. 

Prepare a report for presentation at the next Monster lab meeting about this dataset.

Rubric:

Exploration (summary statistics (the m's), univariate graphs, multivariate graphs)
Regression (Models explored, diagnostics completed, final model choice, justification)

Please include a concluding paragraph (or two) about the implications of your findings.

```{r}
#load up necessary packages
library(HistData)
library(car)
library(stargazer)
```

```{r}
#read in dataset
library(readr)
White_wines <- read_csv("~/Documents/Big Data Class/N741 Data Wrangling/InClass2.1/Regression-1/White_wines.csv")

#Why did I have to do below?
names(White_wines) <- make.names(names(White_wines))
```

```{r}
# see what is in the dataset
summary(White_wines)
```

```{r}
library(Rcmdr)
```


```{r}
# divide the dataset into a training and a testing set based on a random uniform number on fixed seed
set.seed(02132017)
White_wines$group <- runif(length(White_wines$pH), min = 0, max = 1)
summary(White_wines)
White_wines.train <- subset(White_wines, group <= 0.90)
White_wines.test <- subset(White_wines, group > 0.90)

```


```{r}
scatterplotMatrix(~quality+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines.train)

scatterplotMatrix(~quality+pH+density+free.sulfur.dioxide+chlorides,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines.train)

scatterplotMatrix(~quality+citric.acid+fixed.acidity+alcohol,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines.train)
```

Free sulfur dioxide, fixed acidity, and density seem to have abnormal residual distributions, so I logged them below.

```{r}
White_wines.train$free.sulf.diox.log <- with(White_wines.train, 
  log2(free.sulfur.dioxide))
```

```{r}
White_wines.train$fixed.acidity.log <- with(White_wines.train, log2(fixed.acidity))
```

```{r}
White_wines.train$density.log <- with(White_wines.train, log2(density))
```

```{r}
scatterplotMatrix(~quality+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines.train)

scatterplotMatrix(~quality+pH+density.log+free.sulf.diox.log+chlorides,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines.train)

scatterplotMatrix(~quality+citric.acid+fixed.acidity.log+alcohol,
   reg.line=FALSE, smooth=TRUE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'histogram', data=White_wines.train)
```

Those transformations help a little bit.

```{r}
cor(White_wines.train[,c("quality","alcohol","chlorides","citric.acid","density.log","fixed.acidity.log","free.sulf.diox.log","pH","residual.sugar","sulphates","total.sulfur.dioxide","volatile.acidity")], use="complete")
```
The variable most correlated with quality is alcohol, so I will use that as my primary independent variable. Density is highly correlated with alcohol (r=-.78) and residual sugar (r=.84),and free sulfur dioxide less so with total sulfur dioxide (r=.59), so I will consider whether these variables all need to be in the model.


Regression with training data.

```{r}
RegModel.1 <- 
  lm(quality~alcohol+chlorides+citric.acid+density.log+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines.train)
summary(RegModel.1)
```
**Interpretation:**
Density standard error is very high.

```{r}
RegModel.1.2 <- 
  lm(quality~alcohol+chlorides+citric.acid+density+fixed.acidity+free.sulfur.dioxide+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines.train)
summary(RegModel.1.2)
```

The logged variables seem to matter.

```{r}
RegModel.2 <- 
  lm(quality~alcohol+chlorides+citric.acid+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines.train)
summary(RegModel.2)
```
**Interpretation:** for each increase in parent height of 1 inch, the child height increases by 0.65 inches.
With density dropped, there are no anomolous standard errors. R-squared decreases slightly, but not enough to be practically significant.


```{r}
RegModel.3 <- 
  lm(quality~alcohol+chlorides+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines.train)
summary(RegModel.3)
```

Let's see what happens without density and then citric acid.

```{r results="asis"}
# compare the results of the two regression models
stargazer(RegModel.1,RegModel.2,RegModel.3,title="Comparison of 2 Regression outputs",type="html",align=TRUE)
```

```{r}
# diagnostics using residual plots
residualPlots(RegModel.1)
residualPlots(RegModel.2)
residualPlots(RegModel.3)
```

```{r}
White_wines.train$tot.sulf.diox.log <- with(White_wines.train, 
  log2(total.sulfur.dioxide))
```

```{r}
RegModel.4 <- 
  lm(quality~alcohol+chlorides+fixed.acidity.log+free.sulf.diox.log+pH+residual.sugar+sulphates+tot.sulf.diox.log+volatile.acidity,
   data=White_wines.train)
summary(RegModel.4)

residualPlots(RegModel.4)
```

```{r}
RegModel.5 <- 
  lm(quality~alcohol+chlorides+fixed.acidity.log+pH+residual.sugar+sulphates+volatile.acidity,
   data=White_wines.train)
summary(RegModel.5)

residualPlots(RegModel.5)
```

Lose too much R-squared. Stick with 4

```{r}

#added variable plots
avPlots(RegModel.4, id.n=3, id.cex=0.7)
#id.n - identify n most influential observations
#id.cex - controls the size of the dot
```

```{r}

# run the qq-plot
qqPlot(RegModel.4, id.n=3)
# here, id.n identifies the n observations with the largest residuals in absolute value
```

Are there any outliers?

```{r}
 
#run Bonferroni test for outliers
outlierTest(RegModel.4)
```

Are there any points that are of high influence?

```{r}

#identify highly influential points
influenceIndexPlot(RegModel.4, id.n=3)
```

NB. If there are points that are a) outliers AND b) highly influential, these have potential to change the inference. You should consider removing them.

How do we make heads or tails out of the plots above? One way is with an influence plot.
```{r}
#make influence plot
influencePlot(RegModel.4, id.n=3)
```

Another diagnostic is to test for heteroskedasticity (i.e., the variance of the error term is not constant).

```{r}
#test for heteroskedasticity
ncvTest(RegModel.4)
```

We also want to look for multicollinearity, that is are some of our independent variables highly correlated. We do this by looking at the Variance Inflation Factor (VIF). A GVIF > 4 suggests collinearity.

```{r}
vif(RegModel.4)
```